{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-06 thesis/education.tex
dave copyedit description of bugs to be case studied

{\bf Thread exit use-after-free.}
During {\tt thr\_exit()}, a thread must arrange for its own stack memory allocation to be freed.
A common student pitfall is to allow a {\tt thr\_join()}ing thread to free said stack space
while {\tt thr\_exit()} is still executing userspace C code (which inevitably accesses the stack),
or even for {\tt thr\_exit()} to do it itself.
In the former case, threads must interleave specifically to exhibit a use-after-free;
in the latter case, the use-after-free will be deterministic (i.e., present in all interleavings).
Besides Landslide, however,
the students have no Valgrind-like heap debugging tool which would report a bug immediately upon any illegal heap access;
without such, a third thread would have to race to recycle the memory for a new thread stack
and conflict with the old exiting thread
before any problem could be detected.
% TODO: mark which of the tej bugs were deterministic vs race

The Landslide-friendly test {\tt thr\_exit\_join} is most likely to expose this bug, and
the official rubric penalties for it are
{\tt :exit\_stack\_minor -2} and
{\tt :exit\_stack\_race -4} (for the former case),
and
{\tt :uses\_freed\_stack -4} (for the latter).
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.

{\bf Paraguay.}
Named after a certain Paraguayan 15-410 TA from years gone by who originally discovered it,
this refers to a condition variable bug in which a thread which sleeps on two different condition variables
can spuriously wake up early from the later {\tt cond\_wait()}.
The precise reasoning why many common na\"ive student implementations are susceptible to this bug,
as well as the 3 major ways of fixing it,
are one of 15-410 staff's closely-guarded secrets.
Suffice to say that the test {\tt paraguay} invokes a specially-made Pathos {\tt misbehave} mode
which biases thread scheduling towards the specific interleaving required to exhibit the bug
(Landslide, of course, replaces this {\tt misbehav}iour with model checking).
% TODO: i feel tihs sentence should go somewhere else, like after analysis, not in description
Hence, despite being a subtle concurrency bug,
the official course test suite is likely to expose it,
so comparing how many Landslide users and non-users submitted this bug in particular
speaks more to the impact of Landslide's preemption traces
in helping students diagnose a bug that a ``stress'' test could already find.

{\tt paraguay} is itself the Landslide-friendly test for its eponymous bug, and
the official rubric penalty for it is
{\tt :paraguay -2}.
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.

{\bf R/W-lock downgrade.}
In addition to the standard R/W-lock interface,
P2 requires students to implement {\tt rwlock\_downgrade()}:
called with the lock held in write mode,
the caller adopts the reader role instead,
allowing other waiting reader threads to proceed simultaneously,
all while allowing no waiting writers to access in between.

The Landslide-friendly test {\tt rwlock\_downgrade\_read\_test}
checks that readers are allowed simultaneous access after a downgrade, and
the official rubric penalties related to this bug are
{\tt :downgrade\_no\_signal -1},
%{\tt :unsafe\_rwlock\_downgrade -2}, % don't think this is right; not relevant anyway (overlaps)
{\tt :downgrade\_partial\_signal -1}, % not sure if this means smth that dgr would expose
{\tt :downgrade\_\allowbreak{}lies -3}, and
{\tt :downgrade\_deadlock -4}.
Whether each student's submitted P2 solved or suffered from this bug
was determined by checking their grade files for any of the above-listed penalties
(which were assessed by the TAs as a result of their manual inspection, rather than mine).

>>>>

{\bf Thread-exit use-after-free.}
%During {\tt thr\_exit()}, a thread must arrange for its own stack memory allocation to be freed.
After a thread finishes exiting, the memory allocated for its stack must be reclaimed for other uses.
A common student pitfall is to allow a {\tt thr\_join()}ing thread to free said stack space
while {\tt thr\_exit()} is still executing userspace C code (which typically accesses the stack),
or even for {\tt thr\_exit()} to free it itself.
In the former case, threads must interleave specifically to exhibit a use-after-free;
in the latter case, the use-after-free will be deterministic (i.e., present in all interleavings).
Other than Landslide, however,
the students have no Valgrind-like heap debugging tool which would report a bug immediately upon any illegal heap access.
This means that a subsequent {\tt thr\_create()} invocation would need to race to recycle the memory for a new thread stack
and conflict with the old exiting thread
before any problem could be detected.
% TODO: mark which of the tej bugs were deterministic vs race
The Landslide-friendly test {\tt thr\_exit\_join} is most likely to expose this bug.
%, and
%the official rubric penalties for it are
%{\tt :exit\_stack\_minor -2} and
%{\tt :exit\_stack\_race -4} (for the former case),
%and
%{\tt :uses\_freed\_stack -4} (for the latter).
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.

{\bf Paraguay.}
Named after Ivan Jager, the Paraguayan 15-410 TA from 2004-2006 who originally discovered it,
this refers to a condition-variable bug in which a thread which sequentially sleeps on two different condition variables
can spuriously wake up early from the second {\tt cond\_wait()}.
The precise reasoning why many common na\"ive student implementations are susceptible to this bug,
as well as the 3 major ways of fixing it,
are one of 15-410 staff's closely-guarded secrets (see \sect{\ref{sec:410-secrecy}}).
Suffice it to say that the test {\tt paraguay} invokes a specially-made Pathos {\tt misbehave} mode
which biases thread scheduling towards the specific interleaving required to exhibit the bug
(Landslide, of course, replaces this {\tt misbehav}iour with model checking).
% TODO: i feel tihs sentence should go somewhere else, like after analysis, not in description
Hence, despite being a subtle concurrency bug,
the official course test suite is likely to expose it,
so comparing how many Landslide users and non-users submitted this bug in particular
speaks more to the impact of Landslide's preemption traces
in helping students diagnose a bug that a ``stress'' test could already find.
%
{\tt paraguay} is itself the Landslide-friendly test for its eponymous bug.
%, and
%the official rubric penalty for it is
%{\tt :paraguay -2}.
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.

{\bf R/W-lock downgrade.}
In addition to the standard R/W-lock interface,
P2 requires students to implement {\tt rwlock\_downgrade()}:
called with the lock held in write mode,
the caller adopts the reader role instead,
allowing other waiting reader threads to proceed simultaneously,
all while allowing no waiting writers to access in between.
%
The Landslide-friendly test {\tt rwlock\_downgrade\_read\_test}
checks that readers are allowed simultaneous access after a downgrade with no possibility for deadlock.
Rather than one specific bug, this refers to any of several failures that can arise during a downgrade.
%, and
%the official rubric penalties related to this bug are
%{\tt :downgrade\_no\_signal -1},
%%{\tt :unsafe\_rwlock\_downgrade -2}, % don't think this is right; not relevant anyway (overlaps)
%{\tt :downgrade\_partial\_signal -1}, % not sure if this means smth that dgr would expose
%{\tt :downgrade\_\allowbreak{}lies -3}, and
%{\tt :downgrade\_deadlock -4}.
Whether each student's submitted P2 solved or suffered from these bugs
was determined by checking their grade files for any of the above-listed penalties
(which were assessed by the TAs as a result of their manual inspection, rather than mine).

### 2018-07-06 thesis/landslide.tex
dave copyedit description of bugs to be case studied

Many of the supplied P2 test cases invoke the {\tt misbehave()} system call
with a mysterious argument (usually {\tt BGND\_BRWN >> FGND\_CYAN})
before the creation of any child threads.
The use of terminal color code constants is of course a red herring of obfuscation,
as the true nature of the Pathos reference kernel's {\tt misbehave} modes
is a closely-guarded secret among 15-410 course staff.
The mode in question causes the reference kernel to prioritize scheduling the child thread over the parent
whenever {\tt thread\_fork} is called,
and the target thread over the invoking thread
whenever {\tt make\_runnable} is called,
which are necessary to allow Landslide to recognize a {\tt yield()} preemption point
and be able to run the newly-runnable thread as soon as possible.

>>>>

Many of the supplied P2 test cases invoke the {\tt misbehave()} system call
with a mysterious argument (usually {\tt BGND\_BRWN >> FGND\_CYAN})
before the creation of any child threads.
The use of terminal color code constants is of course a red herring of obfuscation,
as the true nature of the Pathos reference kernel's {\tt misbehave} modes
is a closely-guarded secret among 15-410 course staff (\sect{\ref{sec:410-secrecy}}).
The mode in question causes the reference kernel to prioritize scheduling the child thread over the parent
whenever {\tt thread\_fork} is called,
and the target thread over the invoking thread
whenever {\tt make\_runnable} is called,
which are necessary to allow Landslide to recognize a {\tt yield()} preemption point
and be able to run the newly-runnable thread as soon as possible.

### 2018-07-06 thesis/overview.tex
pintos testce
dave copyedit description of bugs to be case studied

@@

\section{Glossary}

@@

% concurrency bug
% race condition: "a confusing term which in some circles means concurrency bug and others data race, which we will avoid"
% data race
% atomicity violation (what even is this.)
% thread
% transition
% conflict
% independent: "see conflict"
% state space
% interleaving
% DPOR
% estimation (WBE, RE)
% landslide
% quicksand
% iterative deepening
% MC, stateless MC
% TA: teaching assistant
% P2 (note abotu psu - dosent call it p2, but we call it that for both schools here

>>>>

\subsubsection{Secrecy}
\label{sec:410-secrecy}

The 15-410 course staff is notoriously secretive about the nature of many concurrency bugs
students commonly encounter during P2 and P3.
This is driven by a desire to cause students to find, diagnose, and fix these bugs on their own during the projects,
rather than to be surprised by them afterwards during grading
\cite{de0u-2018}.
%
One such example is the {\tt paraguay} unit test distributed with P2 (\sect{\ref{sec:education-pebbles-tests}}),
which targets a subtle condition-variable bug.
The test uses the {\tt misbehave} system call to target a particular thread interleaving likely to expose the bug
which is otherwise very unlikely to arise in normal execution.
The reference kernel specification \cite{kspec} does not define the {\tt misbehave} modes' behaviours,
as doing so would deprive students of the learning experience of discovering the interleaving in question on their own.
%
In this thesis I will occasionally use intentionally vague phrasing to preserve the mystery of these bugs.

@@

\section{Glossary}
\label{sec:glossary}

@@

% concurrency bug
% race condition: "a confusing term which in some circles means concurrency bug and others data race, which we will avoid"
% data race
% atomicity violation (what even is this.)
% thread
% transition
% conflict
% independent: "see conflict"
% state space
% interleaving
% DPOR
% estimation (WBE, RE)
% landslide
% quicksand
% iterative deepening
% MC, stateless MC
% TA: teaching assistant
% P2 (note abotu psu - dosent call it p2, but we call it that for both schools here
% happens before, all 3 versions - DPOR, LHB, and PHB
% userspace, kernelspace
